---
title: "GAMMs analyses Trombone - Notes vs. vowels (NZE and Tongan)"
author:
  - Jalal Al-Tamimi (Newcastle University)
  - Donald Derrick (University of Canterbury)
  - Matthias Heyne (Boston University)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: yes
  html_document:
    toc: yes
    toc_depth: '6'
---

This notebook provides additional plots to H1 and H2 of the full analysis of the article: Heyne, M., Derrick, D., and Al-Tamimi, J. (under review). "Native language influence on brass instrument performance: An application of generalized additive mixed models (GAMMs) to midsagittal ultrasound images of the tongue". Frontiers Research Topic: Models and Theories of Speech Production. Ed. Adamantios Gafos & Pascal van Lieshout.

This provides various plots for the parametric terms and for inspecting random effects


# Loading packages & custom plotting function

```{r warning=FALSE, message=FALSE, error=FALSE}
load_packages = c("mgcv","itsadug")
# dplyr, rlist, and plotly are required by the custom plotting functions
for(pkg in load_packages){
  eval(bquote(library(.(pkg))))
  if (paste0("package:", pkg) %in% search()){
    cat(paste0("Successfully loaded the ", pkg, " package.\n"))
  }else{
    install.packages(pkg)
    eval(bquote(library(.(pkg))))
    if (paste0("package:", pkg) %in% search()){
      cat(paste0("Successfully loaded the ", pkg, " package.\n"))
    }
  }
}
rm(load_packages, pkg)

```

```{r warning=FALSE, message=FALSE, error=FALSE}
# specify directory to save models and summaries
output_dir = "updated_models"
```

# Description

These plots look at the parametric terms and at the random effects. 

# Uploading latest model

```{r warning=FALSE, message=FALSE, error=FALSE}
Notes.gam.AR.Mod2 = readRDS(paste0(output_dir,"/Notes.gam.AR.Mod2.rds"))

```

# Plotting parametric terms

Parametric terms are our fixed effects and we use these to look at variance in the data (with standard error) and position of the particular note (and intensity) in both languages. It summarises the variation observed. We look at the distribution of notes across the two languages. The x axis now has the Rho (height) values: the lower the value the lower the overall tongue shape is, and vice-versa.

We observe overall that Tongan has a higher tongue position (overall) compared to NZE. Notes are different as well in that (overall) high notes show slightly higher tongue position and less variability; low notes show slightly lower tongue position and more variability


## Bb2

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb2.piano', 'Tongan.Bb2.piano')),
                cond=list(noteIntenInt='Bb2.piano'), rm.ranef=TRUE,
                main="")
```


### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb2.mezzopiano', 'Tongan.Bb2.mezzopiano')),
                cond=list(noteIntenInt='Bb2.mezzopiano'), rm.ranef=TRUE,
                                main="")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb2.mezzoforte', 'Tongan.Bb2.mezzoforte')),
                cond=list(noteIntenInt='Bb2.mezzoforte'), rm.ranef=TRUE,
                                main="")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb2.forte', 'Tongan.Bb2.forte')),
                cond=list(noteIntenInt='Bb2.forte'), rm.ranef=TRUE,
                                main="")
```



## F3

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F3.piano', 'Tongan.F3.piano')),
                cond=list(noteIntenInt='F3.piano'), rm.ranef=TRUE,
                main="")
```


### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F3.mezzopiano', 'Tongan.F3.mezzopiano')),
                cond=list(noteIntenInt='F3.mezzopiano'), rm.ranef=TRUE,
                                main="")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F3.mezzoforte', 'Tongan.F3.mezzoforte')),
                cond=list(noteIntenInt='F3.mezzoforte'), rm.ranef=TRUE,
                                main="")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F3.forte', 'Tongan.F3.forte')),
                cond=list(noteIntenInt='F3.forte'), rm.ranef=TRUE,
                                main="")
```


## Bb3

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb3.piano', 'Tongan.Bb3.piano')),
                cond=list(noteIntenInt='Bb3.piano'), rm.ranef=TRUE,
                main="")
```


### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb3.mezzopiano', 'Tongan.Bb3.mezzopiano')),
                cond=list(noteIntenInt='Bb3.mezzopiano'), rm.ranef=TRUE,
                                main="")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb3.mezzoforte', 'Tongan.Bb3.mezzoforte')),
                cond=list(noteIntenInt='Bb3.mezzoforte'), rm.ranef=TRUE,
                                main="")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.Bb3.forte', 'Tongan.Bb3.forte')),
                cond=list(noteIntenInt='Bb3.forte'), rm.ranef=TRUE,
                                main="")
```



## D4

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.D4.piano', 'Tongan.D4.piano')),
                cond=list(noteIntenInt='D4.piano'), rm.ranef=TRUE,
                main="")
```


### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.D4.mezzopiano', 'Tongan.D4.mezzopiano')),
                cond=list(noteIntenInt='D4.mezzopiano'), rm.ranef=TRUE,
                                main="")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.D4.mezzoforte', 'Tongan.D4.mezzoforte')),
                cond=list(noteIntenInt='D4.mezzoforte'), rm.ranef=TRUE,
                                main="")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.D4.forte', 'Tongan.D4.forte')),
                cond=list(noteIntenInt='D4.forte'), rm.ranef=TRUE,
                                main="")
```


## F4

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F4.piano', 'Tongan.F4.piano')),
                cond=list(noteIntenInt='F4.piano'), rm.ranef=TRUE,
                main="")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F4.mezzoforte', 'Tongan.F4.mezzoforte')),
                cond=list(noteIntenInt='F4.mezzoforte'), rm.ranef=TRUE,
                                main="")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,1))
plot_parametric(Notes.gam.AR.Mod2,  xlim=c(190,280),
                pred=list(langNoteInt.ord=c('NZE.F4.forte', 'Tongan.F4.forte')),
                cond=list(noteIntenInt='F4.forte'), rm.ranef=TRUE,
                                main="")
```



# Inspecting random effects

Within random effects, we look at how variable our speakers are. Recall that this variability is taken into account in our model as all coefficients are adjusted to account for this variation. We look at the spread of variation in the data. Overall, we observe that our Tongan speakers display most of the variation compared to NZE speakers. This seems to match the predictions for H1b as well.


## Bb2

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=40, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb2 piano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=40, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb2 piano - Tongan")
```



### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=45, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb2 mezzopiano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=45, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb2 mezzopiano - Tongan")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=50, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb2 mezzoforte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=50, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb2 mezzoforte - Tongan")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=55, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb2 forte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=55, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb2 forte - Tongan")
```


## F3

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=41, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F3 piano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=41, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F3 piano - Tongan")
```



### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=46, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F3 mezzopiano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=46, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F3 mezzopiano - Tongan")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=51, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F3 mezzoforte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=51, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F3 mezzoforte - Tongan")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=56, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F3 forte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=56, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F3 forte - Tongan")
```


## Bb3

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=42, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb3 piano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=42, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb3 piano - Tongan")
```



### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=47, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb3 mezzopiano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=47, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb3 mezzopiano - Tongan")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=52, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb3 mezzoforte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=52, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb3 mezzoforte - Tongan")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=57, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="Bb3 forte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=57, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="Bb3 forte - Tongan")
```


## D4

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=43, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="D4 piano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=43, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="D4 piano - Tongan")
```



### Mezzopiano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=48, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="D4 mezzopiano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=48, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="D4 mezzopiano - Tongan")
```


### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=53, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="D4 mezzoforte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=53, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="D4 mezzoforte - Tongan")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=58, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="D4 forte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=58, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="D4 forte - Tongan")
```


## F4

### Piano

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=44, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F4 piano - NZE")
inspect_random(Notes.gam.AR.Mod2, select=44, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F4 piano - Tongan")
```



### Mezzoforte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=54, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F4 mezzoforte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=54, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F4 mezzoforte - Tongan")
```


### Forte

```{r warning=FALSE, message=FALSE, error=FALSE}
par(mfrow=c(1,2))
inspect_random(Notes.gam.AR.Mod2, select=59, ylim=c(-60,60), 
               cond=list(subject=c('S1','S3','S5',
                                   'S12', 'S24', 'S25',
                                   'S26', 'S27',  'S29',
                                   'S30')), col=1, xpd=TRUE,
               main="F4 forte - NZE")
inspect_random(Notes.gam.AR.Mod2, select=59, ylim=c(-60,60),
               cond=list(subject=c('S4','S14','S15',
                                   'S16', 'S17', 'S18',
                                   'S19', 'S20',  'S21',
                                   'S22')), col=2, xpd=TRUE, 
               main="F4 forte - Tongan")
```




