---
title: "GAMMs analyses Trombone - Tongan vs English (NZE)"
author:
  - Jalal Al-Tamimi (Newcastle University)
  - Donald Derrick (University of Canterbury)
  - Matthias Heyne (Boston University)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: yes
  html_document:
    toc: yes
    toc_depth: '6'
---

This notebook provides the first half of the full analysis of the article: Heyne, M., Derrick, D., and Al-Tamimi, J. (under review). "Native language influence on brass instrument performance: An application of generalized additive mixed models (GAMMs) to midsagittal ultrasound images of the tongue". Frontiers Research Topic: Models and Theories of Speech Production. Ed. Adamantios Gafos & Pascal van Lieshout.

```{r warning=FALSE, message=FALSE, error=FALSE}
# specify directory to save models and summaries
output_dir = "updated_models"

# specify whether to run models -> if set to false script will attempt to load saved models from output_dir
run_models = FALSE
```

# Loading packages

```{r warning=FALSE, message=FALSE, error=FALSE}
load_packages = c("readr","knitr","ggplot2","mgcv","itsadug","parallel","dplyr","rlist","plotly")
# dplyr, rlist, and plotly are required by the custom plotting functions
for(pkg in load_packages){
  eval(bquote(library(.(pkg))))
  if (paste0("package:", pkg) %in% search()){
    cat(paste0("Successfully loaded the ", pkg, " package.\n"))
  }else{
    install.packages(pkg)
    eval(bquote(library(.(pkg))))
    if (paste0("package:", pkg) %in% search()){
      cat(paste0("Successfully loaded the ", pkg, " package.\n"))
    }
  }
}
rm(load_packages, pkg)

# detect number of cores available for model calculations
ncores = detectCores()
cat(paste0("Number of cores available for model calculations set to ", ncores, "."))
```

# Loading custom plotting function

## plotly_scatterpolar_multiplot function (Matthias Heyne, 2019)

```{r warning=FALSE, message=FALSE, error=FALSE}
# This function plots multiple smoothing splines in the same window
plotly_scatterpolar_multiplot <- function(df, horizontal, vertical, cols2plot, print=TRUE){
  if (length(cols2plot)>2){
    print("ERROR: You specified more than 2 columns of values to plot.")
  }else{
    dat1=df
    df_name=deparse(substitute(df))
    # layout option 1
    if (length(horizontal)==2 & length(vertical)==1){
      # Note, Intensity, Language
      hori1=nrow(unique(select(dat1, horizontal[1])))
      hori2=nrow(unique(select(dat1, horizontal[2])))
      hori=hori1*hori2
      vert=nrow(unique(select(dat1, vertical[1])))
      dat1=select(dat1, c(horizontal[1],horizontal[2],vertical[1],cols2plot[1],cols2plot[2]))
      dat1=droplevels(dat1)
      var_hori1=levels(dat1[,1])
      var_hori2=levels(dat1[,2])
      var_vert1=levels(dat1[,3])
      
      # set up line types & colors
      ltypes=list("","dash") # match length of hori1
      colors=list("blue","green","orange","red") # match length of hori2
      cat(paste0("Proceeding to assemble a ", hori, "x", vert, " multiplot.\n"))
      cat(paste0("Your plot will show the columns/variables ",horizontal[1]," & ",horizontal[2]," in the horizontal direction and ",vertical[1]," in the vertical direction.\n"))
      cat(paste0(horizontal[1], " will be plotted using the following linestyles: -> "))
      for (n in 1:length(var_hori1)){
        if (n<length(var_hori1)){
          cat(paste0(var_hori1[n], ": ", ltypes[n], " - "))
        }else{
          cat(paste0(var_hori1[n], ": ", ltypes[n], "\n"))
        }
      }
      cat(paste0(horizontal[2], " will be plotted using the following colors: -> "))
      for (n in 1:length(var_hori2)){
        if (n<length(var_hori2)){
          cat(paste0(var_hori2[n], ": ", colors[n], " - "))
        }else{
          cat(paste0(var_hori2[n], ": ", colors[n], "\n"))
        }
      }
      rm(n)
      cat(paste0(vertical[1], " will be shown in the vertical direction from ", var_vert1[1], " (bottom) to ", var_vert1[length(var_vert1)], " (top).\n"))

      # assemble layout options for all subplots
      # plot_specs set as default
      plot_specs = list(sector=c(20,160), radialaxis=list(angle=90, range=c(0,max(dat1$rho_uncut_z)), tickfont=list(size=2)), 
                        angularaxis=list(thetaunit='radians', direction="clockwise", rotation=0, tickfont=list(size=4)))
      # set layout options for required number of subplots
      for (i in 1:hori){
        for (j in 1:vert){
          specsX=list.append(plot_specs, domain=list(x=c((i-1)/hori+(1/hori*0.2), i/hori-1/hori*0.2), 
                                                     y=c((j-1)/vert+(1/vert*0.1),j/vert-1/vert*0.1)))
          assign(paste0("sub_plot",((j-1)*hori)+i), specsX)
        }
      }
      rm(i, j, specsX)
      
      # assemble smoothing splines for traces
      for (j in 1:vert){
        # subset data set by vertical
        dat2=dat1[dat1[,3]==var_vert1[j],]
        for (i1 in 1:hori1){
          # subset data set by horizontal[1]
          dat3=dat2[dat2[,1]==var_hori1[i1],]
          for (i2 in 1:hori2){
            # subset data set by horizontal[2]
            dat4=dat3[dat3[,2]==var_hori2[i2],]
            if (!nrow(dat4)==0){
              if ((((j-1)*hori)+((i1-1)*hori2)+i2)==1){
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[i2], dash=ltypes[i1]))
                assign(paste0("trace",((j-1)*hori)+((i1-1)*hori2)+i2),traceX)
              }else if ((((j-1)*hori)+((i1-1)*hori2)+i2)<=hori){
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[i2], dash=ltypes[i1]), subplot=paste0("polar",((j-1)*hori)+((i1-1)*hori2)+i2))
                assign(paste0("trace",((j-1)*hori)+((i1-1)*hori2)+i2),traceX)
              }else{
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[i2], dash=ltypes[i1]), subplot=paste0("polar",((j-1)*hori)+((i1-1)*hori2)+i2), showlegend=FALSE)
                assign(paste0("trace",((j-1)*hori)+((i1-1)*hori2)+i2),traceX)
            }
            }
            }
        }
      }
      rm(j, i1, i2, traceX, dat2, dat3, dat4)

      # plot assembled traces with assembed layout specifications
      p = plot_ly(type='scatterpolar', mode='lines')
      dont_plot=c()
      p = add_trace(p, theta=trace1$theta, r=trace1$r, line=list(color=trace1$line$color[[1]], dash=trace1$line$dash[[1]]))
      for (k in 2:(hori*vert)){
        if (exists(paste0("trace",k))){
          p = add_trace(p, theta=get(paste0("trace",k))$theta, r=get(paste0("trace",k))$r, 
                        subplot=get(paste0("trace",k))$subplot, 
                        line=list(color=get(paste0("trace",k))$line$color[[1]], dash=get(paste0("trace",k))$line$dash[[1]]))
        }else{
          dont_plot=c(dont_plot,k)
        }
      }
      
      # set layout
      layout_comp = capture.output(
        for (l in 1:(hori*vert)){
          if (is.na(match(l, dont_plot))){
            if (l==1){
              cat(paste0("layout(p, polar=sub_plot",l,", "))
            }else if (l<=hori){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else if (l<hori*vert){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else{
              cat(paste0("polar",l,"=sub_plot",l,", showlegend=FALSE)"))
            }
          }
        })
      p; eval(parse(text=layout_comp))
      
    # layout option 2
    }else if (length(horizontal)==1 & length(vertical)==2){
      # Subject, Note, Intensity
      hori=nrow(unique(select(dat1, horizontal[1])))
      vert1=nrow(unique(select(dat1, vertical[1])))
      vert2=nrow(unique(select(dat1, vertical[2])))
      vert=vert1*vert2
      dat1=select(dat1, c(horizontal[1],vertical[1],vertical[2],cols2plot[1],cols2plot[2]))
      # dat1[,1]=horizontal[1]; dat1[,2]=horizontal[2]; dat1[,3]=vertical[1];
      dat1=droplevels(dat1)
      var_hori1=levels(dat1[,1])
      var_vert1=levels(dat1[,2])
      var_vert2=levels(dat1[,3])
      
      # set up line types & colors
      colors=list("blue","green","orange","red","gray") # match length of vert1
      ltypes=list("","dash","dashdot","dot") # match length of vert2
      cat(paste0("Proceeding to assemble a ", hori, "x", vert, " multiplot.\n"))
      cat(paste0("Your plot will show the columns/variables ",vertical[1]," & ",vertical[2]," in the vertical direction and ",horizontal[1]," in the horizontal direction.\n"))
      cat(paste0(vertical[1], " will be plotted using the following colors: -> "))
      for (n in 1:length(var_vert1)){
        if (n<length(var_vert1)){
          cat(paste0(var_vert1[n], ": ", colors[n], " - "))
        }else{
          cat(paste0(var_vert1[n], ": ", colors[n], "\n"))
        }
      }
      cat(paste0(vertical[2], " will be plotted using the following linestyles: -> "))
      for (n in 1:length(var_vert2)){
        if (n<length(var_vert2)){
          cat(paste0(var_vert2[n], ": ", ltypes[n], " - "))
        }else{
          cat(paste0(var_vert2[n], ": ", ltypes[n], "\n"))
        }
      }
      rm(n)
      cat(paste0(horizontal[1], " will be shown in the horizontal direction from ", var_hori1[1], " (left) to ", var_hori1[length(var_hori1)], " (right).\n"))

      # assemble layout options for all subplots
      # plot_specs set as default
      plot_specs = list(sector=c(20,160), radialaxis=list(angle=90, range=c(0,max(dat1$rho_uncut_z)), tickfont=list(size=2)), 
                        angularaxis=list(thetaunit='radians', direction="clockwise", rotation=0, tickfont=list(size=4)))
      # set layout options for required number of subplots
      for (i in 1:hori){
        for (j in 1:vert){
          specsX=list.append(plot_specs, domain=list(x=c((i-1)/hori+(1/hori*0.2), i/hori-1/hori*0.2), 
                                                     y=c((j-1)/vert+(1/vert*0.1),j/vert-1/vert*0.1)))
          assign(paste0("sub_plot",((j-1)*hori)+i), specsX)
        }
      }
      rm(i, j, specsX)
      
      # assemble smoothing splines for traces
      for (i in 1:hori){
        # subset data set by horizontal
        dat2=dat1[dat1[,1]==var_hori1[i],]
        for (j1 in 1:vert1){
          # subset data set by vertical[1]
          dat3=dat2[dat2[,2]==var_vert1[j1],]
          for (j2 in 1:vert2){
            # subset data set by vertical[2]
            dat4=dat3[dat3[,3]==var_vert2[j2],]
            if (!nrow(dat4)==0){
              if (i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)==1){
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[j1], dash=ltypes[j2]))
                assign(paste0("trace", i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)), traceX)
              }else if (i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)<=hori){
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[j1], dash=ltypes[j2]), subplot=paste0("polar",i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)))
                assign(paste0("trace",i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)),traceX)
              }else{
                # assemble trace & assign number
                traceX=list(theta=seq(min(dat4$theta_uncut_z)*180/pi,max(dat4$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat4$theta_uncut_z, dat4$rho_uncut_z),
                                      seq(min(dat4$theta_uncut_z),max(dat4$theta_uncut_z), length=100))$y,
                            line=list(color=colors[j1], dash=ltypes[j2]), subplot=paste0("polar",i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)), showlegend=FALSE)
                assign(paste0("trace",i+((j1-1)*vert)+((j2-1)*hori)+((j1-1)*vert)),traceX)
              }
            }
          }
        }
      }
      rm(i, j1, j2, traceX, dat2, dat3, dat4)
      
      # plot assembled traces with assembed layout specifications
      p = plot_ly(type='scatterpolar', mode='lines')
      dont_plot=c()
      p = add_trace(p, theta=trace1$theta, r=trace1$r, line=list(color=trace1$line$color[[1]], dash=trace1$line$dash[[1]]))
      for (k in 2:(hori*vert)){
        if (exists(paste0("trace",k))){
          p = add_trace(p, theta=get(paste0("trace",k))$theta, r=get(paste0("trace",k))$r, 
                        subplot=get(paste0("trace",k))$subplot, 
                        line=list(color=get(paste0("trace",k))$line$color[[1]], dash=get(paste0("trace",k))$line$dash[[1]]))
        }else{
          dont_plot=c(dont_plot,k)
        }
      }

      # set layout
      layout_comp = capture.output(
        for (l in 1:(hori*vert)){
          if (is.na(match(l, dont_plot))){
            if (l==1){
              cat(paste0("layout(p, polar=sub_plot",l,", "))
            }else if (l<=hori){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else if (l<hori*vert){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else{
              cat(paste0("polar",l,"=sub_plot",l,", showlegend=FALSE)"))
            }
          }
        })
      p; eval(parse(text=layout_comp))
      
    # layout option 3
    }else if (length(horizontal)==1 & length(vertical)==1){
      # Subject, tokenPooled
      hori=nrow(unique(select(dat1, horizontal[1])))
      vert=nrow(unique(select(dat1, vertical[1])))
      dat1=select(dat1, c(horizontal[1],vertical[1],cols2plot[1],cols2plot[2]))
      dat1=droplevels(dat1)
      var_hori1=levels(dat1[,1])
      var_vert1=levels(dat1[,2])
        
      # set up line types & colors
      if (unique(df$native_lg=="Tongan") && vertical=="token"){
        # levels(dfTongan$token)
        colors=list("#D50D0B","#D50D0B","#003380","#003380","#FF7B00","#FF7B00","#009737","#009737","#C20088","#C20088","#191919","#191919","#191919","#191919","#191919")
        ltypes=list("","dash","","dash","","dash","","dash","","dash","","dash","dashdot","dot","dash")
      }else if (unique(df$native_lg=="Tongan") && vertical=="tokenPooled"){
        colors=list("#D50D0B","#003380","#FF7B00","#009737","#C20088","#191919","#191919","#191919","#191919","#191919")
        ltypes=list("","","","","","","dash","dashdot","dot","dash")
      }else if (unique(df$native_lg=="NZE")){
        # levels(dfNZE$token)
        colors=list("#D50D0B","#990000","#0075DC","#E082B4","#003380","#FF7B00","#009737","#00AFC3","#C20088","#8F48B7","#ACB500","#7B4937","#6C6C6C","#191919","#191919","#191919","#191919","#191919")
        ltypes=list("","","","","","","","","","","","","","","dash","dashdot","dot","dash")
      }
      cat(paste0("Proceeding to assemble a ", hori, "x", vert, " multiplot.\n"))
      cat(paste0("Your plot will show the columns/variables ",horizontal[1]," in the horizontal direction and ",vertical[1]," in the vertical direction.\n"))
      cat(paste0(vertical[1], " will be shown in the vertical direction from ", var_vert1[1], " (bottom) to ", var_vert1[length(var_vert1)], " (top).\n"))
      
      # assemble layout options for all subplots
      # plot_specs set as default
      plot_specs = list(sector=c(20,160), radialaxis=list(angle=90, range=c(0,max(dat1$rho_uncut_z)), tickfont=list(size=2)), 
                        angularaxis=list(thetaunit='radians', direction="clockwise", rotation=0, tickfont=list(size=4)))
      # set layout options for required number of subplots
      for (i in 1:hori){
        for (j in 1:vert){
          specsX=list.append(plot_specs, domain=list(x=c((i-1)/hori+(1/hori*0.2), i/hori-1/hori*0.2), 
                                                     y=c((j-1)/vert+(1/vert*0.1),j/vert-1/vert*0.1)))
          assign(paste0("sub_plot",((j-1)*hori)+i), specsX)
        }
      }
      rm(i, j, specsX)
      
      # assemble smoothing splines for traces
      for (i in 1:hori){
        # subset data set by horizontal
        dat2=dat1[dat1[,1]==var_hori1[i],]
        for (j in 1:vert){
          # subset data set by vertical[1]
          dat3=dat2[dat2[,2]==var_vert1[j],]
          if (!nrow(dat3)==0){
            if (i+(j-1)*hori==1){
              # assemble trace & assign number
              traceX=list(theta=seq(min(dat3$theta_uncut_z)*180/pi,max(dat3$theta_uncut_z)*180/pi, length=100), 
                          r=predict(smooth.spline(dat3$theta_uncut_z, dat3$rho_uncut_z),
                                    seq(min(dat3$theta_uncut_z),max(dat3$theta_uncut_z), length=100))$y,
                          line=list(color=colors[j], dash=ltypes[j]))
              assign(paste0("trace", i+(j-1)*hori), traceX)
            }else if (i+(j-1)*hori<=hori){
              # assemble trace & assign number
              traceX=list(theta=seq(min(dat3$theta_uncut_z)*180/pi,max(dat3$theta_uncut_z)*180/pi, length=100), 
                            r=predict(smooth.spline(dat3$theta_uncut_z, dat3$rho_uncut_z),
                                      seq(min(dat3$theta_uncut_z),max(dat3$theta_uncut_z), length=100))$y,
                            line=list(color=colors[j], dash=ltypes[j]), subplot=paste0("polar",i+(j-1)*hori))
              assign(paste0("trace", i+(j-1)*hori), traceX)
            }else{
              # assemble trace & assign number
              traceX=list(theta=seq(min(dat3$theta_uncut_z)*180/pi,max(dat3$theta_uncut_z)*180/pi, length=100), 
                          r=predict(smooth.spline(dat3$theta_uncut_z, dat3$rho_uncut_z),
                                    seq(min(dat3$theta_uncut_z),max(dat3$theta_uncut_z), length=100))$y,
                          line=list(color=colors[j], dash=ltypes[j]), subplot=paste0("polar",i+(j-1)*hori), showlegend=FALSE)
              assign(paste0("trace", i+(j-1)*hori), traceX)
            }
          }
        }
      }
      rm(i, j, traceX, dat2, dat3)
      
      # plot assembled traces with assembed layout specifications
      p = plot_ly(type='scatterpolar', mode='lines')
      dont_plot=c()
      p = add_trace(p, theta=trace1$theta, r=trace1$r, line=list(color=trace1$line$color[[1]], dash=trace1$line$dash[[1]]))
      for (k in 2:(hori*vert)){
        if (exists(paste0("trace",k))){
          p = add_trace(p, theta=get(paste0("trace",k))$theta, r=get(paste0("trace",k))$r, 
                        subplot=get(paste0("trace",k))$subplot, 
                        line=list(color=get(paste0("trace",k))$line$color[[1]], dash=get(paste0("trace",k))$line$dash[[1]]))
        }else{
          dont_plot=c(dont_plot,k)
        }
      }
      
      # set layout
      layout_comp = capture.output(
        for (l in 1:(hori*vert)){
          if (is.na(match(l, dont_plot))){
            if (l==1){
              cat(paste0("layout(p, polar=sub_plot",l,", "))
            }else if (l<=hori){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else if (l<hori*vert){
              cat(paste0("polar",l,"=sub_plot",l,", "))
            }else{
              cat(paste0("polar",l,"=sub_plot",l,", showlegend=FALSE)"))
            }
          }
        })
      p; eval(parse(text=layout_comp))
    }else{
      cat("Sorry, this layout is not yet implemented in the function. Currently the options are either 2 variables shown horizontally and 1 shown vertically or 1 horizontally and 2 vertically.\n")
      cat("Usage: plotly_scatterpolar_multiplot(df, horizontal, vertical, cols2plot, print=TRUE) ->\n where df refers to the data.frame to plot, horizontal & vertical specify the column names to use as grouping variables,\n and cols2plot refers to the 2 columns of values to plot.\n")
      cat("Use the c(x, y) notation to specify multiple colums for horizontal and/or vertical and for the cols2plot columns.\n")
  }
  }
}
```


# Dataset

## Manipulation

```{r warning=FALSE, message=FALSE, error=FALSE}
df <- read.csv("all_data_NZE_Tongan_w_context_checked_7_March_2019_not_cut_checked_27_Apr.csv", sep=',', stringsAsFactors = F)

# remove empty column
df$X = NULL

df$tokenPooled <- factor(df$tokenPooled)
df$subject <- factor(df$subject)
df$sex <- factor(df$sex)
df$native_lg <- factor(df$native_lg)

df$playing_proficiency[df$playing_proficiency == "intermediate"] <- "amateur"
df$playing_proficiency <- factor(df$playing_proficiency, levels = c("amateur","semi-professional","professional"))

df$block <- factor(df$block)
df$point <- as.numeric(df$point)

df$note_intensity <- factor(df$note_intensity, levels = c("piano","mezzopiano","mezzoforte","forte"))
```


## Two new datasets

### Notes

```{r warning=FALSE, message=FALSE, error=FALSE}
dfNotes <- subset(df,df$tokenPooled == "Bb2"|
                    df$tokenPooled == "Bb3"|
                    df$tokenPooled == "D4"|
                    df$tokenPooled == "F3"|
                    df$tokenPooled == "F4"
                  )
dfNotes$tokenPooled <- factor(dfNotes$tokenPooled, levels = c("Bb2","F3","Bb3","D4","F4"))

dfNotes <- subset(dfNotes,dfNotes$note_intensity == "forte"|
                    dfNotes$note_intensity == "mezzoforte"|
                    dfNotes$note_intensity == "mezzopiano"|
                    dfNotes$note_intensity == "piano"
                  )
dfNotes$note_intensity <- factor(dfNotes$note_intensity, levels = c("piano","mezzopiano","mezzoforte","forte"))

levels(dfNotes$tokenPooled)
levels(dfNotes$note_intensity)
levels(dfNotes$playing_proficiency)
```


### Vowels

```{r}
dfVowelsFull <- subset(df,is.na(df$note_intensity))
dfVowelsFull$tokenPooled <- factor(dfVowelsFull$tokenPooled)
levels(dfVowelsFull$tokenPooled)
# creating a new dataset dfVowels that contains comparable vowels
# in the two languages
dfVowels <- subset(dfVowelsFull,dfVowelsFull$tokenPooled == "ɐ"|
                     dfVowelsFull$tokenPooled == "ɐː"|
                     dfVowelsFull$tokenPooled == "a"|
                     dfVowelsFull$tokenPooled == "e"|
                     dfVowelsFull$tokenPooled == "i"|
                     dfVowelsFull$tokenPooled == "iː"|
                     dfVowelsFull$tokenPooled == "ɛ"|
                     dfVowelsFull$tokenPooled == "o"|
                     dfVowelsFull$tokenPooled == "oː"|
                     dfVowelsFull$tokenPooled == "u"|
                     dfVowelsFull$tokenPooled == "ʉː"
                     )
dfVowels$tokenPooled <- factor(dfVowels$tokenPooled)
levels(dfVowels$tokenPooled)
# creating a new variable TokenPooledNew and changing names of
# vowels in NZE to match those in Tongan.
dfVowels$tokenPooledNew <- dfVowels$tokenPooled
dfVowels$tokenPooledNew <- as.character(dfVowels$tokenPooledNew)
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "ɐ"] <- "a"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "ɐː"] <- "a"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "ʉː"] <- "u"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "oː"] <- "o"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "iː"] <- "i"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "e" &
                          dfVowels$native_lg == "NZE"] <- "i"
dfVowels$tokenPooledNew[dfVowels$tokenPooledNew == "ɛ"] <- "e"

dfVowels$tokenPooledNew <- as.factor(dfVowels$tokenPooledNew)
levels(dfVowels$tokenPooledNew)

dfVowels$speech_preceeding_soundNew <- dfVowels$speech_preceeding_sound
dfVowels <- subset(dfVowels, dfVowels$speech_preceeding_soundNew != "ɹ" &
                     dfVowels$speech_preceeding_soundNew != "ʒ" &
                     dfVowels$speech_preceeding_soundNew != "ð" &
                     dfVowels$speech_preceeding_soundNew != "N" &
                     dfVowels$speech_preceeding_soundNew != "o" &
                     dfVowels$speech_preceeding_soundNew != "v" &
                     dfVowels$speech_preceeding_soundNew != "w" &
                     dfVowels$speech_preceeding_soundNew != "z"
                     )
dfVowels$speech_preceeding_soundNew <- factor(dfVowels$speech_preceeding_soundNew)
levels(dfVowels$speech_preceeding_soundNew)

dfVowels$speech_following_soundNew <- dfVowels$speech_following_sound
dfVowels <- subset(dfVowels, dfVowels$speech_following_soundNew != "ə" &
                     dfVowels$speech_following_soundNew != "ɫ" &
                     dfVowels$speech_following_soundNew != "ɹ" &
                     dfVowels$speech_following_soundNew != "ɾ" &
                     dfVowels$speech_following_soundNew != "ʒ" &
                     dfVowels$speech_following_soundNew != "θ" &
                     dfVowels$speech_following_soundNew != "a" &
                     dfVowels$speech_following_soundNew != "b" &
                     dfVowels$speech_following_soundNew != "ð" &
                     dfVowels$speech_following_soundNew != "g" &
                     dfVowels$speech_following_soundNew != "i" &
                     dfVowels$speech_following_soundNew != "o" &
                     dfVowels$speech_following_soundNew != "u" &
                     dfVowels$speech_following_soundNew != "e" &
                     dfVowels$speech_following_soundNew != "w"
                     )
dfVowels$speech_following_soundNew <- factor(dfVowels$speech_following_soundNew)
levels(dfVowels$speech_following_soundNew)
```


## Tables to check structure

```{r warning=FALSE, message=FALSE, error=FALSE}
kable(table(dfNotes$tokenPooled,dfNotes$note_intensity),format="html")
kable(table(dfNotes$tokenPooled,dfNotes$native_lg),format="html")
kable(table(dfNotes$note_intensity,dfNotes$native_lg),format="html")
kable(table(dfNotes$playing_proficiency,dfNotes$native_lg),format="html")
kable(table(dfNotes$age,dfNotes$native_lg),format="html")
kable(table(dfNotes$sex,dfNotes$native_lg),format="html")

kable(table(dfVowels$tokenPooledNew,dfVowels$native_lg),format="html")
kable(table(dfVowels$age,dfVowels$native_lg),format="html")
kable(table(dfVowels$sex,dfVowels$native_lg),format="html")
kable(table(dfVowels$sex,dfVowels$native_lg),format="html")
kable(table(dfVowels$speech_preceeding_soundNew,dfVowels$native_lg),format="html")
kable(table(dfVowels$speech_following_soundNew,dfVowels$native_lg),format="html")
kable(table(dfVowels$speech_preceeding_soundNew,dfVowels$tokenPooledNew),format="html")
kable(table(dfVowels$speech_following_soundNew,dfVowels$tokenPooledNew),format="html")
```


## Visualising the data

### Notes

Before running anything, we start by visualising the data


#### By Note, By Intensity and By Language

```{r warning=FALSE, message=FALSE, error=FALSE, fig.width=20, fig.height=20}
# run multiplot function
plotly_scatterpolar_multiplot(df=dfNotes, horizontal=c("native_lg","note_intensity"), vertical="tokenPooled", cols2plot=c("theta_uncut_z","rho_uncut_z"))
```


#### By Note, By Intensity, and By subject

##### NZE

Let's start with the NZE data. We see that speakers are variable in how they are producing the notes according to their intensities (which is normal).

```{r warning=FALSE, message=FALSE, error=FALSE, fig.width=20, fig.height=20}
dfNotesNZE <- subset(dfNotes, dfNotes$native_lg == "NZE")
dfNotesNZE$subject <- factor(dfNotesNZE$subject)

plotly_scatterpolar_multiplot(df=dfNotesNZE, horizontal="subject", vertical=c("tokenPooled","note_intensity"), cols2plot=c("theta_uncut_z","rho_uncut_z"))
```


##### Tongan

Moving on to the Tongan data, we see again that speakers are variable in how they are producing the notes according to their intensities (which is normal).

```{r warning=FALSE, message=FALSE, error=FALSE, fig.width=20, fig.height=20}
dfNotesTongan <- subset(dfNotes, dfNotes$native_lg == "Tongan")
dfNotesTongan$subject <- factor(dfNotesTongan$subject)

plotly_scatterpolar_multiplot(df=dfNotesTongan, horizontal="subject", vertical=c("tokenPooled","note_intensity"), cols2plot=c("theta_uncut_z","rho_uncut_z"))
```

#### New variables

The two observations on variations by note and intensity (that are within subject) are important. We will create a variable that combines subject, note and intensity. This will allow us later on to use these instead of subject only to model the within subject variation with respect to the two other predictors (note and intensity). We also create a new variable that combines language, note and intensity of the note to be used as a fixed effect and in adjustments. 

```{r warning=FALSE, message=FALSE, error=FALSE}
dfNotes$noteIntenInt <- interaction(dfNotes$tokenPooled, dfNotes$note_intensity)
levels(dfNotes$noteIntenInt)
str(dfNotes$noteIntenInt)
dfNotes$langNoteInt <- interaction(dfNotes$native_lg, 
dfNotes$tokenPooled, dfNotes$note_intensity)
levels(dfNotes$langNoteInt)
str(dfNotes$langNoteInt)
```


# GAMM

## Ordering predictors

We are intersted in the effect of Native Language on the tongue position of musical notes in relation to their intensity. We create three new ordered predictors

```{r warning=FALSE, message=FALSE, error=FALSE}
dfNotes$native_lg.ord <- as.ordered(dfNotes$native_lg)
contrasts(dfNotes$native_lg.ord) <- "contr.treatment"
dfNotes$tokenPooled.ord <- as.ordered(dfNotes$tokenPooled)
contrasts(dfNotes$tokenPooled.ord) <- "contr.treatment"
dfNotes$note_intensity.ord <- as.ordered(dfNotes$note_intensity)
contrasts(dfNotes$note_intensity.ord) <- "contr.treatment"
dfNotes$langNoteInt.ord <- as.ordered(dfNotes$langNoteInt)
contrasts(dfNotes$langNoteInt.ord) <- "contr.treatment"

dfVowels$native_lg.ord <- as.ordered(dfVowels$native_lg)
contrasts(dfVowels$native_lg.ord) <- "contr.treatment"
dfVowels$tokenPooledNew.ord <- as.ordered(dfVowels$tokenPooledNew)
contrasts(dfVowels$tokenPooledNew.ord) <- "contr.treatment"
```


We create a new variable (start) when Point of tongue == 1. Our dataset is already ordered by speaker, by note, by intensity and by points of measurements.

```{r warning=FALSE, message=FALSE, error=FALSE}
dfNotes$start <- dfNotes$points==1
```


## Running model with no random effects

### Model specification

We start by running a model with no random effects. Just to evaluate structure.

```{r warning=FALSE, message=FALSE, error=FALSE}
if (run_models == TRUE){
  mdl.sys.time1 <- system.time(Notes.gam.noAR.noRandom <- bam(rho_uncut_z ~ langNoteInt.ord +
                          ## 1d smooths
                          s(theta_uncut_z, bs="cr", k=10) +
                          ## 1d smooths * factors
                          s(theta_uncut_z, k=10, bs="cr", by=langNoteInt.ord),
                          data=dfNotes,discrete=TRUE,nthreads=ncores))
mdl.sys.time1
 
# # save model so that it can be reloaded later
 saveRDS(Notes.gam.noAR.noRandom, paste0(output_dir,"/Notes.gam.noAR.noRandom.rds"))
 capture.output(summary(Notes.gam.noAR.noRandom), file =
                  paste0(output_dir,"/summary_Notes.gam.noAR.noRandom.txt"))
}else{
# reload model
Notes.gam.noAR.noRandom = readRDS(paste0(output_dir,"/Notes.gam.noAR.noRandom.rds"))
}
```

### Summary

```{r}
summary(Notes.gam.noAR.noRandom)
```


## Models with random effects

Our second model includes random effects for subject. Given the observed variations between the notes produced by speakers and according to intensity in both languages, we examine two models:

1. A model with interaction between language, note and intensity as fixed effects, and by speaker random smooths
2. A model with interaction between language, note and intensity as fixed effects, and by speaker random smooths with adjustments to note and intensity.


### Optimal models

#### Model specification


```{r warning=FALSE, message=FALSE, error=FALSE}
if (run_models == TRUE){
  mdl.sys.time5 <- system.time(Notes.gam.noAR.Mod1 <- bam(rho_uncut_z ~ langNoteInt.ord +
                          ## 1d smooths
                          s(theta_uncut_z, bs="cr", k=10) +
                          ## 1d smooths * factors
                          s(theta_uncut_z, k=10, bs="cr", by=langNoteInt.ord) +
                          s(theta_uncut_z, subject, bs="fs", k=10, m=1),
                          data=dfNotes, discrete=TRUE, nthreads=ncores))
 mdl.sys.time5
 
# save model so that it can be reloaded later
 saveRDS(Notes.gam.noAR.Mod1, paste0(output_dir,"/Notes.gam.noAR.Mod1.rds"))
 capture.output(summary(Notes.gam.noAR.Mod1), file =
                  paste0(output_dir,"/summary_Notes.gam.noAR.Mod1.txt"))
}else{
# reload model
Notes.gam.noAR.Mod1 = readRDS(paste0(output_dir,"/Notes.gam.noAR.Mod1.rds"))
}
## Model 2
if (run_models == TRUE){
 mdl.sys.time6 <- system.time(Notes.gam.noAR.Mod2 <- bam(rho_uncut_z ~ langNoteInt.ord +
                          ## 1d smooths
                          s(theta_uncut_z, bs="cr", k=10) +
                          ## 1d smooths * factors
                          s(theta_uncut_z, k=10, bs="cr", by=langNoteInt.ord) +
                         ## random smooths by subject, note and intensity
                         s(theta_uncut_z, subject, bs="fs", k=10, m=1, by=noteIntenInt),                
                         data=dfNotes, discrete=TRUE, nthreads=ncores))
 mdl.sys.time6
 
  # save model so that it can be reloaded later
 saveRDS(Notes.gam.noAR.Mod2, paste0(output_dir,"/Notes.gam.noAR.Mod2.rds"))
 capture.output(summary(Notes.gam.noAR.Mod2), file = 
                  paste0(output_dir,"/summary_Notes.gam.noAR.Mod2.txt"))
}else{
# reload model
Notes.gam.noAR.Mod2 = readRDS(paste0(output_dir,"/Notes.gam.noAR.Mod2.rds"))
}
```


#### Checking `k`

##### Model 1

```{r warning=FALSE, message=FALSE, error=FALSE}
gam.check(Notes.gam.noAR.Mod1)
```

##### Model 2

```{r warning=FALSE, message=FALSE, error=FALSE}
gam.check(Notes.gam.noAR.Mod2)
```

#### Summary

##### Model 1

```{r warning=FALSE, message=FALSE, error=FALSE}
summary(Notes.gam.noAR.Mod1)
```

##### Model 2

```{r warning=FALSE, message=FALSE, error=FALSE}
summary(Notes.gam.noAR.Mod2)
```


## Model with random effects and AR1 model

So far, our second model `Notes.gam.noAR.Mod2` that takes into account the random effect structure of by speaker, by note and by intensity accounted for 86.1% of the variance in the data. It showed some differences between the two languages in terms of how tongue contours are different depending on the note and its intensity.
We next need to check the autocorrelation in the residuals and acocunt for these.


### Checking ACF

#### ACF full

As we see below, the autocorrelation in the residuals is massive. We need to check whether this is on all predictors or on specific ones.

```{r warning=FALSE, message=FALSE, error=FALSE}
acf_resid(Notes.gam.noAR.Mod2,main = "Average ACF No.AR",cex.lab=1.5,cex.axis=1.5)
```


## Running our final model

We finally run our final model that takes into account the autocorrelations in the risiduals

### Estimating `Rho`

We use the following to get an estimate of the `rho` to be included later on in our model

```{r}
rho_est <- start_value_rho(Notes.gam.noAR.Mod2)
rho_est
```


### Model specification

```{r warning=FALSE, message=FALSE, error=FALSE}
if (run_models == TRUE){
mdl.sys.time6 <-  system.time(Notes.gam.AR.Mod2 <- bam(rho_uncut_z ~ langNoteInt.ord +
                                            ## 1d smooths
                                            s(theta_uncut_z, bs="cr", k=10) +
                                            ## 1d smooths * factors
                                            s(theta_uncut_z, k=10, bs="cr", by=langNoteInt.ord) +
                                            ## random smooths by subject, note and intensity
                                            s(theta_uncut_z, subject, bs="fs", k=10, m=1,
                                              by=noteIntenInt), 
                                      data=dfNotes,AR.start=dfNotes$start, 
                                      rho=rho_est, discrete=TRUE, nthreads=ncores))
  mdl.sys.time6
 
 # save model so that it can be reloaded later
 saveRDS(Notes.gam.AR.Mod2, paste0(output_dir,"/Notes.gam.AR.Mod2.rds"))
 capture.output(summary(Notes.gam.AR.Mod2), file =
                  paste0(output_dir,"/summary_Notes.gam.AR.Mod2.txt"))
}else{
# reload model
Notes.gam.AR.Mod2 = readRDS(paste0(output_dir,"/Notes.gam.AR.Mod2.rds"))
}
```


### Checking ACF

#### ACF full

```{r warning=FALSE, message=FALSE, error=FALSE}
acf_resid(Notes.gam.AR.Mod2,main = "Average ACF AR", cex.lab=1.5, cex.axis=1.5)
```


### Summary

```{r warning=FALSE, message=FALSE, error=FALSE}
summary(Notes.gam.AR.Mod2)
```


